Kinetic Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm

### 2022-02-03 2022/PAVT/PAVT2022.tex
fix doi

As experimental data, information is taken from the article \cite{cao2019}, which represent a change in the concentrations of reaction components over time at different temperatures; as an example of experimental data at a temperature of 276.2 K, the data in the table \ref{table1} are presented.

\begin{table}
\label{table1}
\caption{Experimental Data}
\begin{center}
\begin{tabular}{ccccccc}
\hline
time, min & 1 & 2 & 5 & 10 & 15 & 20 \\
\hline\rule{0pt}{12pt}
DMH & 0.12 & 0.11 & 0.1	& 0.1 &	0.095 &	0.09  \\
TMP & 0.54 & 0.65 & 0.69 & 0.69 & 0.7 & 0.705 \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}

@@

Step 1. Renumber set of preimages of the trial points 
\[
X_k = \left\{x^1, x^2,...,x^{k+p} \right\},
\]
containing all the preimages, which the trials either have been completed or are being performed at in the increasing order (by the lower index), so that
\[
0=x_1<x_2<...<x_{x+p}=1.
\]
Step 2. Compute the values 
\begin{gather*}
M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},\\
M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},\\
M=\max\{M_1,M_2\},
\end{gather*}
where $z_i=\varphi(y(x_i))$ if $x_i \notin I_k, \; 1\leq i \leq k+p$. The values $z_i$ at the points $x_i \in I_k$ are undefined since the trials at the points $x_i \in I_k$ haven't been completed yet. If the value of $M$ equals 0, then set $M=1$.

@@

First, the problem of searching for activation energies and pre-exponential multipliers of all reaction stages was solved. The number of optimized parameters is 30, which is two parameters for each of the fifteen reaction stages. For activation energies, the search range was set to $0 \leq E_i \leq 100$ kJ/mol, and $0 \leq E_i \leq 10^{12}$ for pre-exponential multipliers based on physicochemical considerations. However, based on the stiffness of the system of differential equations (\ref{eq:one}) and (\ref{eq:twelve}), no solution was found in such a wide range, this is due to the too high degree of pre-exponential multipliers included in the Arrhenius equation. Therefore, during the calculations, the value of the upper bound of the pre-exponential multipliers was reduced and the solution was obtained at the upper bounds of $10^5$. The kinetic parameters found are presented in the table~\ref{table_res1}.

\begin{table}
\label{table_res1}
\caption{Calculated rate constants of reaction}
\begin{center}
\begin{tabular}{cccccccc}
\hline
 & $k_1$ & $k_2$ & $k_3$ & $k_4$ & $k_5$ & $k_6$ & $k_7$\\
\hline\rule{0pt}{12pt}
$E$, kJ / mol & 98.60 & 98.22 & 99.07 & 2.29 & 97.61 & 94.30 & 8.67\\
$k^0$ & $1.46\cdot10^3$ & $1.53\cdot10^2$ & $1.03\cdot10^4$ & $1.10$ & $4.99\cdot10^2$  & $1.31\cdot10^2$ & $5.45\cdot10^2$\\
\hline
$k_8$ & $k_9$ & $k_{10}$ & $k_{11}$ & $k_{12}$ & $k_{13}$ & $k_{14}$ & $k_{15}$ \\
\hline\rule{0pt}{12pt}
11.57 & 73.16 & 65.54 & 13.86 & 4.93 & 2.69 & 21.19 & 0.63\\
$2.60\cdot10^2$ & $3.95\cdot10^4$ & $3.24\cdot10^4$ & $1.73\cdot10$ & $3.43$ & $1.78$ & $1.42\cdot10^3$ & $5.57\cdot10$\\[2pt]
\hline
\end{tabular}\end{center}\end{table}

>>>>

As experimental data, information is taken from the article \cite{cao2019}, which represent a change in the concentrations of reaction components over time at different temperatures; as an example of experimental data at a temperature of 276.2 K, the data in Table \ref{table1} are presented.

\begin{table}
\caption{Experimental Data}
\label{table1}
\begin{center}
\begin{tabular}{ccccccc}
\hline
time, min & 1 & 2 & 5 & 10 & 15 & 20 \\
\hline\rule{0pt}{12pt}
DMH & 0.12 & 0.11 & 0.1	& 0.1 &	0.095 &	0.09  \\
TMP & 0.54 & 0.65 & 0.69 & 0.69 & 0.7 & 0.705 \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}

@@

Step 1. Renumber set of preimages of the trial points 
\[
X_k = \left\{x^1, x^2,...,x^{k+p} \right\},
\]
containing all the preimages, which the trials either have been completed or are being performed at in the increasing order (by the lower index), so that
\[
0=x_1<x_2<...<x_{k+p}=1.
\]
Step 2. Compute the values 
\begin{gather*}
M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},\\
M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},\\
M=\max\{M_1,M_2\},
\end{gather*}
where $z_i=\varphi(y(x_i))$ if $x_i \notin I_k, \; 1\leq i \leq k+p$. The values $z_i$ at the points $x_i \in I_k$ are undefined since the trials at the points $x_i \in I_k$ haven't been completed yet. If the value of $M$ equals 0, then set $M=1$.

@@

First, the problem of searching for activation energies and pre-exponential multipliers of all reaction stages was solved. The number of optimized parameters is 30, which is two parameters for each of the fifteen reaction stages. For activation energies, the search range was set to $0 \leq E_i \leq 100$ kJ/mol, and $0 \leq E_i \leq 10^{12}$ for pre-exponential multipliers based on physicochemical considerations. However, based on the stiffness of the system of differential equations (\ref{eq:one}) and (\ref{eq:twelve}), no solution was found in such a wide range, this is due to the too high degree of pre-exponential multipliers included in the Arrhenius equation. Therefore, during the calculations, the value of the upper bound of the pre-exponential multipliers was reduced and the solution was obtained at the upper bounds of $10^5$. The kinetic parameters found are presented in Table~\ref{table_res1}.

\begin{table}
\caption{Calculated rate constants of reaction}
\label{table_res1}
\begin{center}
\begin{tabular}{cccccccc}
\hline
 & $k_1$ & $k_2$ & $k_3$ & $k_4$ & $k_5$ & $k_6$ & $k_7$\\
\hline\rule{0pt}{12pt}
$E$, kJ / mol & 98.60 & 98.22 & 99.07 & 2.29 & 97.61 & 94.30 & 8.67\\
$k^0$ & $1.46\cdot10^3$ & $1.53\cdot10^2$ & $1.03\cdot10^4$ & $1.10$ & $4.99\cdot10^2$  & $1.31\cdot10^2$ & $5.45\cdot10^2$\\
\hline
$k_8$ & $k_9$ & $k_{10}$ & $k_{11}$ & $k_{12}$ & $k_{13}$ & $k_{14}$ & $k_{15}$ \\
\hline\rule{0pt}{12pt}
11.57 & 73.16 & 65.54 & 13.86 & 4.93 & 2.69 & 21.19 & 0.63\\
$2.60\cdot10^2$ & $3.95\cdot10^4$ & $3.24\cdot10^4$ & $1.73\cdot10$ & $3.43$ & $1.78$ & $1.42\cdot10^3$ & $5.57\cdot10$\\[2pt]
\hline
\end{tabular}\end{center}\end{table}
